---
title: "PretermBirthAnalysis"
author: "Alison Paquette"
date: "9/30/2020"
output: html_document
---

# Load Relevant Data
-load the TRN as it will be presented in the R shiny app so that we can ensure
-This has been published as a pre-print
```{r}
TRNModel<-read.csv("~/Dropbox/PlacentalTRNProject/FinalAnalysisOctober2020/FinalResultsfromKhaleesi/TRNForRShinyApp.csv")
TRNModel<-TRNModel[,-1]
TRNModel$targetGene<-as.character(TRNModel$targetGene)
TRNModel$TF<-as.character(TRNModel$TF)

#Load results of miroarray analysis published in BOR Paper

SigGenes<-read.csv("/Volumes/Active/paquette_a/apaquette/ECHOPathways_CRHAnalysis/IntermediateData/SupplementalTable2_ForTRN.csv")

rownames(SigGenes)<-as.character(SigGenes$Gene)

load("/Users/apaque/Dropbox/PlacentalTRNProject/FinalAnalysisOctober2020/FinalResultsfromKhaleesi/Node&Edge.RData")
```


# Enrichment Test Function####

Perform Fishers exact test to see if Gene list is enriched for specifif TFs.
Reference: https://www.pathwaycommons.org/guide/primers/statistics/fishers_exact_test/
```{r}
EnrichFisher<-function(NSigGenes,NTotalGenes,SigInGroup,GroupSize){

  a=SigInGroup ## Equivalent to X; Table 3. In the TFgroup and DE
  b= GroupSize-SigInGroup #. Equivalent to m-X: In the TFGroup but NOT DE
  c=NSigGenes-SigInGroup  # Equivalent to K-X; DE but not in the group
  n=NTotalGenes-GroupSize #n in contigencyTable: not in group
  d=n-c
  mat<-(matrix(c(a,c,b,d),nrow=2))
  mat
  sum(mat)==NTotalGenes
  fisher.test(mat,alternative="greater")$p.value #default is 2 sided. We need this to be greater rather than less to say that these genes are enriched
}
```

# Generate TF Enrichment Table
-This is basically what the RShiny function is doing
```{r}



TFTable<-as.data.frame(table(TRNModel$TF))
colnames(TFTable)<-c("TF","NTargets")
rownames(TFTable)<-TFTable$TF
TFTable$TF<-as.character(TFTable$TF)
TFTable$SigGenes<-NA

  for(i in 1:dim(TFTable)[1])
  {
    x<-subset(TRNModel,TF==as.character(TFTable$TF[i])) # Pull out the TF of interest
    dim(x)[1]==TFTable$NTargets[i] #This should be equal
    TFTable$SigGenes[i]<-length(intersect(as.character(x$targetGene),SigGenes$Gene)) #How many in the input gene list match this TF?
  }



  TFTable$FisherTest<-NA
  # Perform Enrichment Test using Fishers Exact Function
  for(i in 1:length(TFTable$TF)){
    TFTable$FisherTest[i]<-EnrichFisher(NSigGenes=as.numeric(length(SigGenes$Gene)),NTotalGenes=7712,SigInGroup =TFTable$SigGenes[i],GroupSize=TFTable$NTargets[i])
  }

AllTFTable<-TFTable
# Filter By Preset Thresholds

MinSigGenes=3
#ModuleMax

#TFTable<-subset(TFTable,NTargets>=ModuleMin)
#TFTable<-subset(TFTable,NTargets<ModuleMax)
TFTable<-subset(TFTable,SigGenes>=MinSigGenes)
TFTable$FisherAdjustP<-p.adjust(TFTable$FisherTest,method="BH",n=dim(TFTable)[1])

EnrichedTFTable<-subset(TFTable,FisherAdjustP<0.05)

write.csv(EnrichedTFTable,file="/Volumes/Active/paquette_a/apaquette/ECHOPathways_CRHAnalysis/TRNAnalysis11222023.csv")
```

#Pull out Enriched TFs and get calculations for paper
```{r}
EnrichedTFs<-subset(TFTable,FisherAdjustP<0.05)$TF

#1. How many total Genes in List?
GenesIncluded<-NULL
for(i in 1:length(EnrichedTFs)){
  TargetGenes<-as.character(subset(TRNModel,TF==EnrichedTFs[i])$targetGene)
  TargetGenes<-intersect(TargetGenes,SigGenes$Gene)
  GenesIncluded<-union(TargetGenes,GenesIncluded)
}


length(GenesIncluded)/length(SigGenes$Gene)

Table4<-TFTable[EnrichedTFs,]

Table4$Genes<-NA

for(i in 1:length(EnrichedTFs)){
    TargetGenes<-as.character(subset(TRNModel,TF==EnrichedTFs[i])$targetGene)
    TargetGenes<-intersect(TargetGenes,SigGenes$Gene)
    Table4$Genes[i]<-paste(TargetGenes,collapse=" ")
}

```

# Which CRH genes were ALSO TFs?
```{r}
rownames(AllTFTable)<-AllTFTable$TF
PTBTFs<-TFTable[intersect(SigGenes$Gene,TFTable$TF),]

PTBTFs<-merge(PTBTFs,SigGenes,by='row.names',all.x=T,all.y=F)

PTBTFs$DEGenes<-NA
for(i in 1:length(PTBTFs$TF)){
    TargetGenes<-as.character(subset(TRNModel,TF==PTBTFs$TF[i])$targetGene)
    TargetGenes<-intersect(TargetGenes,SigGenes$Gene)
    PTBTFs$DEGenes[i]<-paste(TargetGenes,collapse=",")
}
write.table(PTBTFs,file="~/Dropbox/PlacentalTRNProject/Manuscript and Docs/TRNManuscript/July2022Submission_TRN/BioRXiv Submission/ScienceAdvancesResponsetoReviewers/ResponsetoReviewrs3_September2023/NewAnalyses/PretermBirthTFs10112023.csv")


#Total Number of Preterm Birth Genes

DE_PTBGenes<-NA
for(i in 1:length(PTBTFs$TF)){
    TargetGenes<-as.character(subset(TRNModel,TF==PTBTFs$TF[i])$targetGene)
    TargetGenes<-intersect(TargetGenes,SigGenes$Gene)
    DE_PTBGenes<-union(TargetGenes,DE_PTBGenes)
}

length(DE_PTBGenes)/length(SigGenes$Gene)
```

